---
title: "Recreating John Snow's map in R"
author: "Your Name"
date: '`r Sys.Date()`'
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: false
    theme: united
    highlight: textmate
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Packages

```{r}
#| echo: false
#| warning: false
#| message: false

library(tidyverse)
library(sf)
library(HistData)

```

# Load Data 

Lets load the `Snow.deaths` data from the `HistData` package

```{r}
snow_deaths <- HistData::Snow.deaths

snow_deaths |> class()

snow_deaths |> head()
```


Let us now convert this `data.frame` into an `sf` object using the `st_as_sf()` function. 

```{r}
snow_deaths_sf <- snow_deaths |> 
  st_as_sf(coords = c('x', 'y'), crs = 4326)

```

Note the additional arguments such as `coords` and `crs`. Here we use the `crs` of 4326. 

Let us now check the class of the newly created `snow_deaths_sf` object.


```{r}
snow_deaths_sf |> class()
```

Note that the `sf` object is also a `data.frame`.

Let us look at the structure of the `sf` object.

```{r}
snow_deaths_sf
```

Let us now visualise the deaths using the `geom_sf()` function.

```{r}
snow_deaths_sf |> 
  ggplot() +
  geom_sf() 

ggplot() +
  geom_sf(data = snow_deaths_sf)

```

Let us now add color to the deaths using the `color` argument.

```{r}
snow_deaths_sf |> 
  ggplot() +
  geom_sf(color = 'red') 
```


Now, we can add additional layers such as the water pumps and the streets of Soho, London.


```{r}
snow_streets <- HistData::Snow.streets

snow_streets |> class()

snow_streets |> head()
```

We may have to convert the lat long coordinates into lines. 


```{r}
snow_streets_sf <- snow_streets |> 
  st_as_sf(coords = c('x', 'y'), crs = 4326) |> 
  group_by(street) |> 
  summarize(n = mean(n)) |> 
  st_cast('LINESTRING')

snow_streets_sf |> 
  slice(400:480)



```

Let us now look at the `sf` object. 

```{r}
snow_streets_sf
```

Note the geometry type is now a `sfc_LINESTRING`.


Let us plot the streets now.

```{r}
snow_streets_sf |> 
  ggplot() + 
  geom_sf()

```


Let us add the streets of Soho, London as a new layer in our previously created deaths map.

```{r}
ggplot() +
  geom_sf(data = snow_streets_sf) +
  geom_sf(data = snow_deaths_sf, color = 'red', alpha = 0.2, size = 3)
```

We can now add a new layer of water pumps on the above map.


```{r eval=TRUE, include=TRUE}
snow_pumps_sf <- HistData::Snow.pumps |> 
  st_as_sf(coords = c('x', 'y'), crs = 4326)

ggplot() +
  geom_sf(data = snow_streets_sf) +
  geom_sf(data = snow_deaths_sf, color = 'red', alpha = 0.2, size = 3) +
  geom_sf(data = snow_pumps_sf , shape = 22, size = 4, fill = 'blue', color = 'blue')  +
  geom_sf_label(data = snow_pumps_sf, aes(label = label), nudge_x = 0.025, nudge_y = -0.5) # Add labels
```




