---
title: "Mapmaking Exercise"
author: "Group Activity - R4PublicHealth Workshop"
date: "2022-12-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Load Packages
library(sf)
library(tidyverse)
library(here)


# Load Spatial file - District level boundary file with population estimates 
# projected for 2020
kerala_district_sf <- read_rds(here("spatial_files", "kerala_district_sf.rds"))


# Load COVID-19 Data for Kerala by different waves
kerala_covid_cases_df <- read_rds(here("spatial_files", "kerala_covid_cases_df.rds"))

```


# Explore non-spatial data

```{r}
kerala_covid_cases_df |> glimpse()

```

# Create a Total Population column

```{r}
# Lets create a total population column
kerala_district_sf <- kerala_district_sf |> 
  mutate(top_pop = sum(pop_2020, na.rm = T))
```


# Visualise the spatial files

```{r}
kerala_district_sf |> 
  ggplot() +
  geom_sf()
```


# Create a choropleth map based on the population of the district

```{r}
kerala_district_sf |> 
  ggplot() +
  geom_sf(aes(fill = pop_2020)) +
  scale_fill_distiller(palette = "Spectral")

```

# Join both the spatial and non-spatial data

```{r}
# Step 1: Check if the district names are given correctly
# Check if the district names in both the datasets are the same
(kerala_covid_cases_df |> distinct(District) |> pull(District)) %in% (kerala_district_sf |> distinct(distname) |> pull(distname))

# Step 2: Reshape your data
kerala_covid_cases_df_reshaped <- kerala_covid_cases_df |> 
  pivot_wider(names_from = Wave, values_from = c(Cases, Deaths))

# Step 3: Identify the common variable in both the datasets (here it is `distname` and `District`)
kerala_district_joined_sf <- left_join(kerala_district_sf, kerala_covid_cases_df_reshaped, by = c("distname" = "District"))



```

# ASSIGNMENTS

### 1. Create a new variable named estimating the standarized incidence rate. 
(SIR can be estimated by SIR = `cases / ((total_cases / total_pop) * pop)` )


```{r}
joined_df <- kerala_covid_cases_df |> 
  left_join(kerala_district_sf |> st_drop_geometry(),
            by = c("District" = "distname")) |> 
  group_by(Wave) |> 
  mutate(SIR = Cases / ((sum(Cases)/ top_pop) * pop_2020)) |> 
  ungroup()

kerala_district_sf <- kerala_district_sf |> left_join(joined_df |> select(District, SIR, Wave) |> rename(distname = District))
 

kerala_district_sf |> 
  ggplot() + 
  geom_sf(aes(fill = SIR))  +
  scale_fill_distiller(palette = "Spectral") +
  facet_wrap(~Wave)



```


### 2. Create a new variable named estimating the standarized motality rate. 
(SMR can be estimated by SMR = `deaths / ((total_deaths / total_pop) * pop)` )

```{r}
# Write the code here


```


### 3. Plot a choropleth map of Cases for each of the wave

```{r}
# Write the code here


```



### 4. Plot a choropleth map of Deaths for each of the wave

```{r}
# Write the code here


```



### 5. Plot a choropleth map of SIR for each of the wave

```{r}
# Write the code here


```


### 6. Plot a choropleth map of SMR for each of the wave

```{r}
# Write the code here


```






